\documentclass[11pt]{article}

\usepackage{color}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks]{hyperref}

\def\N{{\mathbb N}}
\def\NN{{\mathcal N}}
\def\R{{\mathbb R}}
\def\E{{\mathbb E}}
\def\rank{{\mathrm{rank}}}
\def\tr{{\mathrm{trace}}}
\def\P{{\mathrm{Prob}}}
\def\sign{{\mathrm{sign}}}
\def\diag{{\mathrm{diag}}}

\setlength{\oddsidemargin}{0.25 in}
\setlength{\evensidemargin}{-0.25 in}
\setlength{\topmargin}{-0.6 in}
\setlength{\textwidth}{6.5 in}
\setlength{\textheight}{8.5 in}
\setlength{\headsep}{0.75 in}
\setlength{\parindent}{0.25 in}
\setlength{\parskip}{0.1 in}

\newcommand{\lecture}[4]{
   \pagestyle{myheadings}
   \thispagestyle{plain}
   \newpage
   \setcounter{page}{1}
   \setcounter{section}{0}
   \noindent
   \begin{center}
   \framebox{
      \vbox{\vspace{2mm}
    \hbox to 6.28in { {\bf A Mathematical Introduction to Data Analysis \hfill #4} }
       \vspace{6mm}
       \hbox to 6.28in { {\Large \hfill #1  \hfill} }
       \vspace{6mm}
       \hbox to 6.28in { {\it Instructor: #2\hfill #3} }
      \vspace{2mm}}
   }
   \end{center}
   \markboth{#1}{#1}
   \vspace*{4mm}
}


\begin{document}

\lecture{Final Project}{Yuan Yao}{Due: midnight Sunday 21 May, 2017}{28 April, 2017}

%The problem below marked by $^*$ is optional with bonus credits. % For the experimental problem, include the source codes which are runnable under standard settings. 
%
%\begin{enumerate}
%
%\item {\em Manifold Learning}: The following codes by Todd Wittman contain major manifold learning algorithms talked on class.
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/Spring2011/matlab/mani.m}
%
%Precisely, eight algorithms are implemented in the codes: MDS, PCA, ISOMAP, LLE, Hessian Eigenmap, Laplacian Eigenmap, Diffusion Map, and LTSA. 
%The following nine examples are given to compare these methods,
%\begin{enumerate}
%\item Swiss roll;
%\item Swiss hole;
%\item Corner Planes;
%\item Punctured Sphere;
%\item Twin Peaks;
%\item 3D Clusters;
%\item Toroidal Helix;
%\item Gaussian;
%\item Occluded Disks.
%\end{enumerate}
%Run the codes for each of the nine examples, and analyze the phenomena you observed. 
%
%\end{enumerate}

%\newpage


\section{Final Project Requirement}

This project aims to exercise the tools in the class based on the real world datasets. But you are encouraged to explore the datasets with any techniques you feel reasonable. In the below, we list some candidate datasets for your reference. 

\begin{enumerate}
\item Pick up ONE (or more if you like) favorite dataset below to work. If you would like to work on a different problem outside the candidates we proposed, please email course instructor about your proposal.  
\item Team work: we encourage you to form small team, up to THREE persons per group, to work on the same problem. Each team just submit ONE report, \emph{with a clear remark on each person's contribution}. The report can be in the format of a \emph{technical report within 8 pages}, e.g. NIPS conference style 
\begin{center}
\url{https://nips.cc/Conferences/2016/PaperInformation/StyleFiles} 
\end{center}
or of a \emph{poster}, e.g. 
\begin{center}\url{http://math.stanford.edu/~yuany/publications/poster_CleaveBioCPH2017_ForReview.pptx}
\end{center}
%A sample poster file with PKU logo can be found at \\ 
%\url{http://math.stanford.edu/~yuany/course/reference/poster_v5.pdf} \\
%whose source LATEX codes can be downloaded at \\
%\url{http://math.stanford.edu/~yuany/course/reference/pkuposter.zip} 
\item In the report, (1) design or raise your scientific problems (a good problem is sometimes more important than solving it); (2) show your main results with a careful analysis supporting the results toward answering your problems. Remember: scientific analysis and reasoning are more important than merely the performance results. Supplementary material such as source codes (and/or personal datasets) may be submitted through email as a zip file, or as an appendix if it is not large.    
\item Submit your report by email or paper version no later than the deadline, to the following address (\href{mailto:datascience.hw@gmail.com}{datascience.hw@gmail.com}) with Title: \underline{Math 6380: Final}. % (\href{mailto:datascience\_hw@126.com}{datascience\_hw@126.com}). 
\end{enumerate}

\section{Ranking with Pairwise Comparisons}

\subsection{Drug Sensitivity Ranking in Kaggle in class contest}

This contest is announced at

 \url{http://https://inclass.kaggle.com/c/drugsensitivity-3}
 
\noindent with the invitation link at \url{https://kaggle.com/join/math6380project3}. This is a URL that can be shared around and re-used. Anyone who visits this link (and satisfies any other invitation rules that have been set) will be able to participate in the competition.

Given cell line $k$ with genetic feature vector $X(k)$ and pairwise sensitivity comparison of drug $i$ and $j$ on $k$, $y(k,i,j)$, a basic generalized linear model can be 
\begin{equation} Prob( y(k,i,j) = 1) = F(\beta_0(i) - \beta_0(j) + \left< X(k), \beta_1(i) - \beta_1(j)\right>). 
\end{equation}
In fact when $F=Id$, it leads to the following linear regression problem: 
\begin{eqnarray}
& \min_{\beta_0,\beta_1 \in \R^{n}} & \frac{1}{2} \sum_{(k,i,j)} ( y(k,i,j) -  (\beta_0(i) - \beta_0(j) + \left< X(k), \beta_1(i) - \beta_1(j)\right>))^2 + P_\lambda(\beta)
\end{eqnarray}
For example, sparsity assumption on $\beta_1$ can be associated with LASSO penalty $P_\lambda(\beta)=\lambda \| \beta\|_1$. 

Moreover, when $F(x) = e^x / (e^x + e^{-x})$ with binary response $y\in \{\pm1\}$, it leads to the logistic regression, e.g. using the Maximum Likelihood principle one can find an estimate by
\begin{eqnarray}
& \min_{\beta_0,\beta_1 \in \R^{n}} &  \sum_{(k,i,j)}  \log(1+ \exp(-y(k,i,j) f((k,i,j),\beta) )) + P_\lambda(\beta)
\end{eqnarray}
where $f((k,i,j),\beta)= \beta_0(i) - \beta_0(j) + \left< X(k), \beta_1(i) - \beta_1(j)\right>$.

A summary of generalized linear models in pairwise ranking problem, when $\beta_1$ is not presented, can be found in the following paper and references therein

\url{http://math.stanford.edu/~yuany/publications/TMM12-final.pdf}

\subsection{WorldCollege Ranking}
WorldCollege
dataset is composed of 261 colleges. Using the Allourideas
crowdsourcing platform, a total of 340 distinct
annotators from various countries (e.g., USA, Canada,
Spain, France, Japan) are shown randomly with pairs of these colleges, and asked to decide which of the two universities
is more attractive to attend. The following dataset contains a
total of 8,823 pairwise comparisons.

\url{http://math.stanford.edu/~yuany/course/data/college.csv}

This dataset has been used for various studies, e.g. Qianqian Xu, Jiechao Xiong, Xiaochun Cao, and Yuan Yao. False Discovery Rate Control and Statistical Quality Assessment of Annotators in Crowdsourced Ranking, ICML 2016, in \url{http://arxiv.org/pdf/1605.05860v1.pdf}

\subsection{Human Age Ranking}


The following dataset is kindly provided by Qianqian Xu, CAS, for the exploration on class. 

The dataset is contained in the following zip file.

\url{http://math.stanford.edu/~yuany/course/data/age.zip}

\noindent where you may find
\begin{enumerate}
\item \texttt{readme.txt}: description of data
\item \texttt{Agedata.mat}: data file collected
\item \texttt{Groundtruth.mat}: Groundtruth
\item \texttt{30 images.zip}: 30 human age images
\end{enumerate}

The basic problem is to \emph{predict the human age}, using all the information collected so far. A simple sub-problem is rank aggregation of ages from pairwise comparisons. If you are interested, you can try some generalized linear models (Qianqian Xu, Qingming Huang, Tingting Jiang, Bowei Yan, Weisi Lin, and Yuan Yao. HodgeRank on Random Graphs for Subjective Video Quality Assessment. IEEE Transactions on Multimedia, 14(3):844-857, 2012, \url{http://math.stanford.edu/~yuany/publications/TMM12-final.pdf}) on this dataset, such as uniform model, Bradley-Terry model, Thurstone-Mosteller model, and Angular transform model. Compare maximum likelihood estimators and least square ones. 

\section{Identification of Raphael's paintings from the forgeries}

The following data, provided by Prof. Yang WANG from HKUST,

\url{https://drive.google.com/folderview?id=0B-yDtwSjhaSCZ2FqN3AxQ3NJNTA&usp=sharing}

\noindent contains a 28 digital paintings of Raphael or forgeries. Note that there are both jpeg and tiff files, so be careful with the bit depth in digitization. The following file

\url{https://docs.google.com/document/d/1tMaaSIrYwNFZZ2cEJdx1DfFscIfERd5Dp2U7K1ekjTI/edit}

\noindent contains the labels of such paintings, which are 
\begin{enumerate}
\item[1] Maybe Raphael - Disputed
\item[2] Raphael
\item[3] Raphael
\item[4] Raphael
\item[5] Raphael
\item[6] Raphael
\item[7] Maybe Raphael - Disputed
\item[8] Raphael
\item[9] Raphael
\item[10] Maybe Raphael - Disputed
\item[11] Not Raphael
\item[12] Not Raphael
\item[13] Not Raphael
\item[14] Not Raphael
\item[15] Not Raphael
\item[16] Not Raphael
\item[17] Not Raphael
\item[18] Not Raphael
\item[19] Not Raphael
\item[20] My Drawing (Raphael?)
\item[21] Raphael
\item[22] Raphael
\item[23] Maybe Raphael - Disputed
\item[24] Raphael
\item[25] Maybe Raphael - Disputed
\item[26] Maybe Raphael - Disputed
\item[27] Raphael
\item[28] Raphael
\end{enumerate}
Can you exploit the known Raphael vs. Not Raphael data to predict the identity of those 6 disputed paintings (maybe Raphael)? The following student poster report seems a good exploration

\url{http://math.stanford.edu/~yuany/course/2015.fall/poster/Raphael_LI\%2CYue_1300010601.pdf}

The following paper by Haixia Liu, Raymond Chan, and me studies Van Gogh's paintings which might be a reference for you:

\url{http://dx.doi.org/10.1016/j.acha.2015.11.005}


\section{Co-appearance data in novels: Dream of Red Mansion and Journey to the West}

A 374-by-475 binary matrix of character-event can be found at the course website, in .XLS, .CSV, .RData, and .MAT formats. For example the RData format is found at

\url{http://math.stanford.edu/~yuany/course/data/dream.RData} 

\noindent with a readme file:

\url{http://math.stanford.edu/~yuany/course/data/dream.Rd}

\noindent as well as the .txt file which is readable by R command {\tt read.table()},

\url{http://math.stanford.edu/~yuany/course/data/HongLouMeng374.txt}

\url{http://math.stanford.edu/~yuany/course/data/readme.m}

Thanks to Ms. WAN, Mengting, who helps clean the data and kindly shares her BS thesis for your reference
 
\url{http://math.stanford.edu/~yuany/report/WANMengTing2013_HLM.pdf}

%Among various choices of analysis, with this data matrix $X$, you may form a weighted graph $W=X * X'$, pursue PCA of $X$, and sparse SVD of $X$ etc. As an example, here is a project presentation by LI, Liying which gives an analysis of A Journey to the West (by Chen-En Wu) based on PCA, for the class Mathematical Introduction to Data Science in Fall 2012 where you may find more interesting approaches.
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/reference/LiyingLI_Xiyouji2012_slides.pdf}

Moreover you may find a similar matrix of 302-by-408 for the Journey to the West (by Chen-En Wu) at:

\url{http://math.stanford.edu/~yuany/course/data/west.RData}

\noindent whose matlab format is saved at

\url{http://math.stanford.edu/~yuany/course/data/xiyouji.mat}

%%%%%%

\section{Jiashun Jin's data on Coauthorship and Citation Networks for Statisticians}
Thanks to Prof. Jiashun Jin at CMU, who provides his collection of citation and coauthor data for statisticians. The data set covers all papers between 2003 and the first quarter of 2012 from the Annals of Statistics, Journal of the American Statistical Association, Biometrika and Journal of the Royal Statistical Society Series B. The paper corrections and errata are not included. There are 3607 authors and 3248 papers in total. The zipped data file (14M) can be found at 

\url{http://math.stanford.edu/~yuany/course/data/jiashun/Jiashun.zip}

\noindent with an explanation file

\url{http://math.stanford.edu/~yuany/course/data/jiashun/ReadMe.txt}

With the aid of Mr. LI, Xiao, a subset consisting 35 COPSS award winners (\url{https://en.wikipedia.org/wiki/COPSS_Presidents\%27_Award}) up to 2015, is contained in the following file

\url{http://math.stanford.edu/~yuany/course/data/copss.txt} 

\noindent An example was given in the following article, A Tutorial of Libra: R Package of Linearized Bregman Algorithms in High Dimensional Statistics, downloaded at

\url{http://math.stanford.edu/~yuany/course/reference/Libra_Tutorial_springer.pdf}

The citation of this dataset is: \emph{P. Ji and J. Jin. Coauthorship and citation networks for statisticians. Ann. Appl. Stat. Volume 10, Number 4 (2016), 1779-1812}, (\url{http://projecteuclid.org/current/euclid.aoas})

\section{NIPS paper datasets}
NIPS is one of the major machine learning conferences. The following datasets collect NIPS papers:

\subsection{NIPS papers (1987-2016)} The following website: 

\url{https://www.kaggle.com/benhamner/nips-papers}

\noindent collects titles, authors, abstracts, and extracted text for all NIPS papers during 1987-2016. In particular the file {\texttt{paper\_authors.csv}} contains a sparse matrix of paper coauthors. 

\subsection{NIPS words (1987-2015)} The following website:

\url{https://archive.ics.uci.edu/ml/datasets/NIPS+Conference+Papers+1987-2015}

\noindent collects the distribution of words in the full text of the NIPS conference papers published from 1987 to 2015. The dataset is in the form of a 11463 x 5812 matrix of word counts, containing 11463 words and 5811 NIPS conference papers (the first column contains the list of words). Each column contains the number of times each word appears in the corresponding document. The names of the columns give information about each document and its timestamp in the following format: {\texttt{Xyear\_paperID}}. 


%\section{Drug Efficacy Data}
%
%Thanks to Prof. Xianting Ding at Shanghai Jiao Tong University and Prof. Chih-Ming Ho from University of California at Los Angeles, we have the following datasets on combinatorial drug efficacy.
%
%The first dataset consists of two experiments, all with the same 4 drugs in cell lines for attacking leukemia, with 256 experiments of combinatorial drug dosage at 4 levels. The response is the therapeutic window measuring the efficacy with a trade-off by toxicity. 
% 
% \url{http://math.stanford.edu/~yuany/course/data/Ding_4drugs.xlsx}
%
%\noindent whose drugs are explained in 
%
%\url{http://math.stanford.edu/~yuany/course/data/Ding_4drugs_readme.pdf}
%
%Can you find a good prediction of drug response efficacy using those combinatorial dosage levels? It was suggested that quadratic polynomials at logarithmic dosage levels are good models in personalized medicine, e.g. the following cover paper in Science \emph{Translation Medicine}:
%
%\url{http://stm.sciencemag.org/content/8/333/333ra49}
%
%\noindent with a sample 14 drug efficacy at level 2 experiment data in liver transplant: 
%
%\url{http://math.stanford.edu/yuany/course/data/TB-FSC-03A-data.xlsx}

\section{Drug Sensitivity Datasets}

\subsection{One Drug data by Prof. Jiguang WANG@HKUST}
The following dataset, provided by Prof. Jiguang WANG and Dr. Biaobin JIANG, contains 642 cancer cell line samples in response to one drug, Afatinib, with 60 binary features as gene mutation status. Column 1 is the sample ID, column 2 is the cell line names, column 3 is the real valued response as drug sensitivity measured by logarithmic IC50, and all the remaining 60 columns are the binary features/predictors. IC50 values are the dosage amounts of the drug such that after a period of time (say 24 hours) the experimental cancer cell lines are killed by 50\%. Therefore the lower is the IC50, the more sensitive is the cancer line to the drug. Usually negative logarithmic IC50s are called \emph{sensitive} while positive logarithmic IC50s are called \emph{resistant}.
\begin{center}
\url{http://math.stanford.edu/~yuany/course/data/OneDrug/OneDrug_train.csv}
\end{center}

A random subset of 100 cell line responses are withdrawn to the test sample. The basic problem is to predict the responses of these test samples based on their binary predictors/features. A sample prediction file is provided in 
\begin{center}
\url{http://math.stanford.edu/~yuany/course/data/OneDrug/OneDrug_submission.csv}
\end{center}

A Kaggle in-class contest is in the following website. 
\begin{center}
\url{https://inclass.kaggle.com/c/drugsensitivity-2}
\end{center}
The invitation link is: \url{https://kaggle.com/join/math6380project2}. This is a URL that can be shared around and re-used.
Anyone who visits this link (and satisfies any other invitation rules that have been set)
will be able to participate in the competition.
%The following dataset is kindly provided by Cleave Co. Ltd. USA, for the exploration on class. {\textbf{Please keep its use only in this class and any publication will be subject to the approval of Cleave.}}
%
%The dataset is contained in the following zip file (73M).
%
%\url{http://math.stanford.edu/~yuany/course/data/cleave.zip}
%
%\noindent where you may find
%\begin{enumerate}
%\item \texttt{data explanation.pptx}: description of data in pptx
%\item \texttt{data for Yuan Yao.xlsx}: data file
%\item \texttt{Gene set collection 1 for Yuan Yao.txt}: gene set collection
%\item \texttt{Gene set collection 2 for Yuan Yao.txt}: gene set collection
%\item \texttt{reference}: a folder contains a survey paper on 40+ machine learning algorithms as well as some source codes -- \emph{Nature Biotechnology 32, 1202--1212 (2014)} (\url{http://www.nature.com/nbt/journal/v32/n12/full/nbt.2877.html})
%\end{enumerate}
%
%The basic problem is to predict the drug response \texttt{IC50 within 72 hours}, using all the information collected so far, introduced by Ms. Lijing Wang with slides
%
%\url{http://math.stanford.edu/~yuany/course/2016.spring/cleave_lijing.pdf}
%
%\noindent as well as our CPH'2017 poster
%
%\url{http://math.stanford.edu/~yuany/publications/poster_CleaveBioCPH2017_ForReview.pdf}
%
%\noindent where the crucial discovery is that recursive variable selection by LASSO is more effective than one-stage LASSO. 


\subsection{Combinatorial Drug 20 Efficacy Data by Prof. Ed Chow@NUS}
The following dataset, provided by Prof. Ed K. CHOW at NUS, contains 120 cancer cell line samples in response to a combination of 20 drugs in 4 dosage levels. Column 1 is the sample ID, column 2 through column 21 list the discrete dosage levels of 20 drugs, column 22 is the real valued response as \emph{viability} measured by difference between normalized cells and cancer cells. The higher is the viability, the more effective is the drug combination. 
\begin{center}
\url{http://math.stanford.edu/~yuany/course/data/Drug20/DrugEfficacy_train.csv}
\end{center}
A random subset of 20 cell line responses are withdrawn to the test sample. The basic problem is to predict the responses of these test samples based on their discrete levels of combinatorial drugs. Due to the expensive experiments in combinatorial drug efficacy measurements, the less features are in your models, the better. A sample prediction file is provided in 
\begin{center}
\url{http://math.stanford.edu/~yuany/course/data/Drug20/DrugEfficacy_submission.csv}
\end{center}
A Kaggle in-class contest is setup in the following website. 
\begin{center}
\url{https://inclass.kaggle.com/c/combodrug20}
\end{center}
The invitation link is: \url{https://kaggle.com/join/math6380edchow}. This is a URL that can be shared around and re-used.
Anyone who visits this link (and satisfies any other invitation rules that have been set) will be able to participate in the competition.

\subsection{Cleave Dataset}
The following dataset is kindly provided by Cleave Co. Ltd. USA, for the exploration on class. {\textbf{Please keep its use only in this class and any publication will be subject to the approval of Cleave.}}

The dataset is contained in the following zip file (73M).

\url{http://math.stanford.edu/~yuany/course/data/cleave.zip}

\noindent where you may find
\begin{enumerate}
\item \texttt{data explanation.pptx}: description of data in pptx
\item \texttt{data for Yuan Yao.xlsx}: data file
\item \texttt{Gene set collection 1 for Yuan Yao.txt}: gene set collection
\item \texttt{Gene set collection 2 for Yuan Yao.txt}: gene set collection
\item \texttt{reference}: a folder contains a survey paper on 40+ machine learning algorithms as well as some source codes -- \emph{Nature Biotechnology 32, 1202--1212 (2014)} (\url{http://www.nature.com/nbt/journal/v32/n12/full/nbt.2877.html})
\end{enumerate}

The basic problem is to predict the drug response \texttt{IC50 within 72 hours}, using all the information collected so far, introduced by Ms. Lijing Wang with slides

\url{http://math.stanford.edu/~yuany/course/2016.spring/cleave_lijing.pdf}

\noindent as well as our CPH'2017 poster

\url{http://math.stanford.edu/~yuany/publications/poster_CleaveBioCPH2017_ForReview.pdf}

\noindent where the crucial discovery is that recursive variable selection by LASSO is more effective than one-stage LASSO. 

In the mini-project 1, we launched a Kaggle in-class contest with this dataset at the following website,

\url{https://inclass.kaggle.com/c/drugsensitivity}

\noindent which is closed right now. But you may still see the leader board and submit one result for evaluations. 

%Thanks to Prof. Xianting Ding at Shanghai Jiao Tong University and Prof. Chih-Ming Ho from University of California at Los Angeles, we have the following datasets on combinatorial drug efficacy.
%
%The first dataset consists of two experiments, all with the same 4 drugs in cell lines for attacking leukemia, with 256 experiments of combinatorial drug dosage at 4 levels. The response is the therapeutic window measuring the efficacy with a trade-off by toxicity. 
% 
% \url{http://math.stanford.edu/~yuany/course/data/Ding_4drugs.xlsx}
%
%\noindent whose drugs are explained in 
%
%\url{http://math.stanford.edu/~yuany/course/data/Ding_4drugs_readme.pdf}
%
%Can you find a good prediction of drug response efficacy using those combinatorial dosage levels? It was suggested that quadratic polynomials at logarithmic dosage levels are good models in personalized medicine, e.g. the following cover paper in Science \emph{Translation Medicine}:
%
%\url{http://stm.sciencemag.org/content/8/333/333ra49}
%
%\noindent with a sample 14 drug efficacy at level 2 experiment data in liver transplant: 
%
%\url{http://math.stanford.edu/yuany/course/data/TB-FSC-03A-data.xlsx}


%\subsection{The Characters in A Dream of Red Mansion} 
%
%A 376-by-475 matrix of character-event can be found at the course website, in .XLS, .CSV, and .MAT formats. For example the Matlab format is found at
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/hongloumeng/hongloumeng376.mat} 
%
%\noindent with a readme file:
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/hongloumeng/readme.m}
%
%Thanks to Ms. WAN, Mengting (now at UIUC), an update of data matrix consisting 374 characters (two of 376 are repeated) which is readable by R read.table() can be found at 
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/hongloumeng/HongLouMeng374.txt}
%
%\noindent She also kindly shares her BS thesis for your reference
% 
% \url{http://www.math.pku.edu.cn/teachers/yaoy/reference/WANMengTing2013_HLM.pdf}
%
%% Among various choices of analysis, with this data matrix $X$, you may form a weighted graph $W=X * X'$, pursue PCA of $X$. 
%
%\subsection{A Journal to the West} On course website, you may also find the link to this dataset with a 302-by-408 matrix, whose matlab format is saved at
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/Fall2011/xiyouji/xiyouji.mat}
%
%For your reference, here is a project presentation by Mr. LI, Liying (at PKU) which gives an analysis based on PCA
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/reference/LiyingLI_Xiyouji2012_slides.pdf}
%

\section{Heart PCI Operation Effect Prediction}

The following data, provided by Dr. Jinwen Wang at Anzhen Hospital in Beijing, 

\url{http://math.stanford.edu/~yuany/course/data/heartData_20140401.xlsx}

\noindent contains 2581 patients with 73 measurements (inputs) as well as a response variable indicating if after the heart operation there is a null-reflux state. This is a classification problem, with a challenge from the large amount of missing values. Sheet 3 and 4 in the file contains some explanation of the data and variables. 

The problems are listed here:
\begin{enumerate}
\item The inputs (covariates) are of three kinds, measurements upon check-in, measurements before PCI operation, and measurements in PCI operations. For doctors, it is desired to find a prediction model based on measurements before the operation (including check-in). Sheet 2 in the file contains only such measurements.
\subitem The following two reports by LV, Yuan and LI, Xiao, respectively, might be interesting to you:

\url{http://math.stanford.edu/~yuany/course/reference/MSThesis.LvYuan.pdf} 

\url{http://arxiv.org/abs/1511.04656} 

\item It is also an interesting problem how to predict the effect based on all measurements, with lots of missing values. Sheet 1 contains the full measurements. There are some good work by previous students, which are listed here for your reference: 
%\subitem The following two reports by LU, Yu and WANG, Qing, are probably inspiring to you.
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/reference/LuYu_201303_BigHeart.pdf} 
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/reference/WangQing_201303_BigHeart.pdf} 

\subitem The following report by MIAO, Wang and LI, Yanfang, pioneers in missing value treatment. 

\url{http://math.stanford.edu/~yuany/course/reference/MiaoLi2013S_project01.pdf}

\end{enumerate} 

Since we did not separate a subset as test data, you may use $k$-fold cross validation (e.g. $k=5$) to evaluate your models. In the mini-project 2, you may start from the prediction with all measurements, playing with various ways of missing value fill-in and feature selections. An interpretable model is preferred!

%\emph{In the final project, it is desired to take only those measurements upon check-in to predict the probability of non-reflux (non-reflow) after PCI operations. An interpretable model adds a big value! You may compare with your first warm-up project to show your improvements.} 


\section{Finance Data}
The following data contains 1258-by-452 matrix with closed prices of 452 stocks in SNP'500 for workdays in 4 years.

\url{http://math.stanford.edu/~yuany/course/data/snp452-data.mat} 

\noindent or in R: 

\url{http://math.stanford.edu/~yuany/course/data/snp500.Rda}

%You may use PCA to explore the `invisible hands' of markets.

\section{Hand-written Digits} The website 

\url{http://www-stat.stanford.edu/\~tibs/ElemStatLearn/datasets/zip.digits/}

\noindent contains images of 10 handwritten digits (`$0$',...,`9');

\section{Animal Sleeping Data} The following data contains animal sleeping hours together with other features: 

\url{http://math.stanford.edu/~yuany/course/data/sleep1.csv}


\section{US Crime Data} The following data contains crime rates in 59 US cities during 1970-1992:

\url{http://math.stanford.edu/~yuany/course/data/crime.zip}

%\section{Air Quality Weibo Data} (courtesy of Prof. Xiaojin Zhu from University of Wisconsin at Madison) 
%You can login my server:
%
%\texttt{ssh einstein@162.105.205.92}
%
%\noindent using the password I provided on class. 
%
%On the read-only folder \texttt{/data/AQweibo/}, the \texttt{AQICityData/} directory contains the Weibo posts, the AQI for 108 cities with (AQI) information during the study period
%from 2013-11-18 to 2013-12-18 (both inclusive); Information for the spatiotempral bin (city,date) is in the directory \texttt{city\_date/}. See \texttt{README.txt} for more information.
%
%

%\section{SNPs Data}
% This dataset contains a data matrix $X\in \R^{p\times n}$ of about $n=650,000$ columns of SNPs (Single Nucleid Polymorphisms) and $p=1064$ rows of peoples around the world. Each element is of three choices, $0$ (for `AA'), $1$ (for `AC'), $2$ (for `CC'), and some missing values marked by $9$. 
%
%\url{http://math.stanford.edu/~yuany/course/ceph_hgdp_minor_code_XNA.txt.zip}
%
%\noindent which is big (151MB in zip and 2GB original txt). Moreover, the following file contains the region where each people comes from, as well as two variables {\texttt{ind1}} and{\texttt{ind2}} such that $X({\texttt{ind1}},{\texttt{ind2}})$ removes all missing values. 
%
%\url{http://math.stanford.edu/~yuany/course/data/HGDP_region.mat}
%
%Some results by PCA can be found in the following paper, Supplementary Information. 
%
%\url{http://www.sciencemag.org/content/319/5866/1100.abstract}
%
%%Attention: this last dataset is relatively big with about 2GB size. 
%
%You can login my server:
%
%\texttt{ssh einstein@162.105.205.92}
%
%\noindent using the password I provided on class. On the read only folder \texttt{/data/snp/}, you will find all the data in both .txt and .mat (\texttt{data.mat, HGDP\_region.mat, readme.m}).



%\subsection{Bird Flu Dataset} (courtesy of Steve Smale and Cissy) This dataset 162 H5N1 (bird flu) virus sequences discovered around the world:
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_seq162.txt} 
%
%Locations of such virus discovered are reported with latitude and longitude coordinates on the globe:
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_latgrat.txt} 
%
%Pairwise geodesic distances between these 162 sites are constructed as  
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_geodist.txt}
%
%A kernel-induced $l_2$-distances between 162 virus sequences are given in 
%
%\url{http://www.math.pku.edu.cn/teachers/yaoy/data/birdflu_l2dist.txt}
\end{document}


